### 2013-03-19 paper/doc/code-cpp-listings.hpp.tex
last changes before arXiv resubmission

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{//listing00}
\PY{c+c1}{// code licensed under the terms of GNU GPL v3}
\PY{c+c1}{// copyright holder: University of Warsaw}
\PY{c+c1}{//listing01}
\PY{k}{typedef} \PY{k+kt}{double} \PY{n}{real\PYZus{}t}\PY{p}{;}
\PY{c+c1}{//listing02}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}blitz}\PY{c+cp}{/}\PY{c+cp}{array.h\PYZgt{}}
\PY{k}{using} \PY{n}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{k}{using} \PY{n}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
\PY{k}{using} \PY{n}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{c+c1}{//listing03}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{define return\PYZus{}macro(expr) \PYZbs{}}
\PY{c+cp}{  -\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
\PY{c+cp}{\PYZob{} return safeToReturn(expr); \PYZcb{} }
\PY{c+c1}{//listing04}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
\PY{k}{struct} \PY{n}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
  \PY{p}{\PYZob{}}   
    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing05}
\PY{k}{struct} \PY{n}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{+}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{p}{;} 
\PY{p}{\PYZcb{}}

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{-}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{;} 
\PY{p}{\PYZcb{}}
\PY{c+c1}{//listing06}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
  \PY{p}{)}\PY{p}{;} 
\PY{p}{\PYZcb{}} 
\PY{c+c1}{//listing07}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)} 
\PY{p}{\PYZob{}}
  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;}

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)} 
\PY{p}{\PYZob{}}
  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;} 
\PY{c+c1}{//listing08}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
  \PY{k+kt}{int} \PY{n}{n}\PY{p}{,} \PY{n}{hlo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
  \PY{n}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
  \PY{n}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{solver}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}\PY{p}{)} \PY{o}{:}
    \PY{n}{hlo}\PY{p}{(}\PY{n}{hlo}\PY{p}{)}\PY{p}{,}
    \PY{n}{n}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{,} 
    \PY{n}{i}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nx}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
    \PY{n}{j}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{ny}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}  
    \PY{n}{bcx}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} 
    \PY{n}{bcy}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{l} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{l} \PY{o}{\PYZlt{}} \PY{l+m+mi}{2}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{l}\PY{p}{)} 
      \PY{n}{psi}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
    \PY{n}{C}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}\PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// accessor methods}
  \PY{n}{arr\PYZus{}t} \PY{n}{state}\PY{p}{(}\PY{p}{)} \PY{p}{\PYZob{}}
    \PY{k}{return} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{.}\PY{n}{reindex}\PY{p}{(}\PY{p}{\PYZob{}}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}

  \PY{n}{arr\PYZus{}t} \PY{n}{courant}\PY{p}{(}\PY{k+kt}{int} \PY{n}{d}\PY{p}{)} 
  \PY{p}{\PYZob{}} 
    \PY{k}{return} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{;} 
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// helper methods invoked by solve()}
  \PY{k}{virtual} \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;}

  \PY{k+kt}{void} \PY{n}{cycle}\PY{p}{(}\PY{p}{)} 
  \PY{p}{\PYZob{}} 
    \PY{n}{n} \PY{o}{=} \PY{p}{(}\PY{n}{n} \PY{o}{+} \PY{l+m+mi}{1}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2} \PY{o}{-} \PY{l+m+mi}{2}\PY{p}{;} 
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// integration logic}
  \PY{k+kt}{void} \PY{n}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{t} \PY{o}{\PYZlt{}} \PY{n}{nt}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{t}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{advop}\PY{p}{(}\PY{p}{)}\PY{p}{;}
      \PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing09}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{cyclic}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{cyclic}\PY{p}{(}
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k+kt}{int} \PY{n}{hlo}
  \PY{p}{)} \PY{o}{:}
    \PY{n}{left\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,}
    \PY{n}{rght\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{-}\PY{n}{hlo}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}  \PY{p}{)}\PY{p}{,}
    \PY{n}{rght\PYZus{}halo}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}  \PY{p}{)}\PY{p}{,}
    \PY{n}{left\PYZus{}edge}\PY{p}{(}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{o}{+}\PY{n}{hlo}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing10}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{T1}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T2}\PY{p}{,} \PY{k}{class} \PY{n+nc}{T3}\PY{o}{\PYZgt{}} 
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{F}\PY{p}{(}
    \PY{k}{const} \PY{n}{T1} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}l}\PY{p}{,} \PY{k}{const} \PY{n}{T2} \PY{o}{\PYZam{}}\PY{n}{psi\PYZus{}r}\PY{p}{,} \PY{k}{const} \PY{n}{T3} \PY{o}{\PYZam{}}\PY{n}{C}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{p}{(}
      \PY{p}{(}\PY{n}{C} \PY{o}{+} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}l} \PY{o}{+} 
      \PY{p}{(}\PY{n}{C} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{)}\PY{p}{)} \PY{o}{*} \PY{n}{psi\PYZus{}r}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{p}{)}
\PY{c+c1}{//listing11}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}  
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{donorcell}\PY{p}{(} 
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{F}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} \PY{o}{-}
    \PY{n}{F}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}   \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,} 
        \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{-}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)}
  \PY{p}{)}
\PY{c+c1}{//listing12}
  \PY{k+kt}{void} \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} \PY{k}{const} \PY{k+kt}{int} \PY{n}{n}\PY{p}{,}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{p}{\PYZob{}} 
    \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}
      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
      \PY{o}{-} \PY{n}{donorcell}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{j}\PY{p}{,} \PY{n}{i}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{c+c1}{//listing13}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{n}{solver\PYZus{}donorcell}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:}
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

  \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} 
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
    \PY{p}{)}\PY{p}{;}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing14}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{nom\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{den\PYZus{}t}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
    \PY{k}{const} \PY{n}{nom\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{nom}\PY{p}{,} \PY{k}{const} \PY{n}{den\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{den}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{where}\PY{p}{(}\PY{n}{den} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{nom} \PY{o}{/} \PY{n}{den}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
  \PY{p}{)} 
\PY{c+c1}{//listing15}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}A}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{,}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} 
  \PY{p}{)} 
\PY{c+c1}{//listing16}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}B}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
   \PY{n}{mpdata\PYZus{}frac}\PY{p}{(}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{-} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{,}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+}
      \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{psi}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{2}
  \PY{p}{)}
\PY{c+c1}{//listing17}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{p}{(}
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{p}{(}
      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{+}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+}
      \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{n}{C}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{o}{-}\PY{n}{h}\PY{p}{)}\PY{p}{)} 
    \PY{p}{)} \PY{o}{/} \PY{l+m+mi}{4}
  \PY{p}{)}
\PY{c+c1}{//listing18}
  \PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
  \PY{k+kr}{inline} \PY{k}{auto} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{p}{(}
    \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{psi}\PY{p}{,} 
    \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,}
    \PY{k}{const} \PY{n}{arrvec\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{C}
  \PY{p}{)} \PY{n}{return\PYZus{}macro}\PY{p}{(}
    \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{p}{(}\PY{l+m+mi}{1} \PY{o}{-} \PY{n}{abs}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}A}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)} 
    \PY{o}{-} \PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{p}{]}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{i}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} 
    \PY{o}{*} \PY{n}{mpdata\PYZus{}C\PYZus{}bar}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{C}\PY{p}{[}\PY{n}{d}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
    \PY{o}{*} \PY{n}{mpdata\PYZus{}B}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{psi}\PY{p}{,} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{)}
  \PY{p}{)} 
\PY{c+c1}{//listing19}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{n\PYZus{}iters}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}mpdata} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{solver\PYZus{}mpdata}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:} 
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
    \PY{n}{im}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,}
    \PY{n}{jm}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{n} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}tmp}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{n}\PY{p}{)}
    \PY{p}{\PYZob{}}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{step} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{step} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}iters}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{step}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{k}{if} \PY{p}{(}\PY{n}{step} \PY{o}{=}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{)} 
        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
        \PY{p}{)}\PY{p}{;}
      \PY{k}{else}
      \PY{p}{\PYZob{}}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{hlo}\PY{p}{)}
        \PY{p}{)}\PY{p}{;}

        \PY{c+c1}{// choosing input/output for antidiff C}
        \PY{k}{const} \PY{n}{arrvec\PYZus{}t} 
          \PY{o}{\PYZam{}}\PY{n}{C\PYZus{}unco} \PY{o}{=} \PY{p}{(}\PY{n}{step} \PY{o}{=}\PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
            \PY{o}{?} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C} 
            \PY{o}{:} \PY{p}{(}\PY{n}{step} \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{)} 
              \PY{o}{?} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}  \PY{c+c1}{// odd steps}
              \PY{o}{:} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{c+c1}{// even steps}
          \PY{o}{\PYZam{}}\PY{n}{C\PYZus{}corr} \PY{o}{=} \PY{p}{(}\PY{n}{step}  \PY{o}{\PYZpc{}} \PY{l+m+mi}{2}\PY{p}{)} 
            \PY{o}{?} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}    \PY{c+c1}{// odd steps}
            \PY{o}{:} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{;}   \PY{c+c1}{// even steps}

        \PY{c+c1}{// calculating the antidiffusive C }
        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{(}\PY{n}{im}\PY{o}{+}\PY{n}{h}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{0}\PY{o}{\PYZgt{}}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{im}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}

        \PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{jm}\PY{o}{+}\PY{n}{h}\PY{p}{)} \PY{o}{=} \PY{n}{mpdata\PYZus{}C\PYZus{}adf}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{[}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{jm}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{n}{C\PYZus{}unco}
        \PY{p}{)}\PY{p}{;}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{C\PYZus{}corr}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{,}\PY{n}{h}\PY{p}{)}\PY{p}{)}\PY{p}{;}

        \PY{c+c1}{// donor-cell step }
        \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
          \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{n}{C\PYZus{}corr}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}
        \PY{p}{)}\PY{p}{;}
      \PY{p}{\PYZcb{}}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing20}
\end{Verbatim}