### 2013-03-19 paper/doc/code-for-test.f.tex

\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{program }\PY{n+nv}{test}
  \PY{k}{use }\PY{n+nv}{solver\PYZus{}mpdata\PYZus{}m}
  \PY{k}{use }\PY{n+nv}{cyclic\PYZus{}m}
  \PY{k}{implicit }\PY{k}{none}

\PY{k}{  }\PY{k}{if} \PY{p}{(}\PY{n+nb}{command\PYZus{}argument\PYZus{}count}\PY{p}{(}\PY{p}{)} \PY{o}{/}\PY{o}{=} \PY{p}{(}\PY{l+m+mi}{9}\PY{p}{)}\PY{p}{)} \PY{k}{then}
\PY{k}{    }\PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"expecting 9 arguments (nx, ny, Cx, Cy, nt, it, f.in, f.out, dec)"}
    \PY{k}{stop }\PY{l+m+mi}{1}
  \PY{k}{end }\PY{k}{if}

\PY{k}{  }\PY{k}{block}
\PY{c}{!listing23}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{mpdata\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{slv}
    \PY{k}{type}\PY{p}{(}\PY{n+nv}{cyclic\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{target} \PY{k+kd}{::} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{,} \PY{n+nv}{nt}\PY{p}{,} \PY{n+nv}{it}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{Cx}\PY{p}{,} \PY{n+nv}{Cy}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{ptr}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
\PY{c}{!listing24}
    \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{dec}

    \PY{n+nv}{nx} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
    \PY{n+nv}{ny} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{2}\PY{p}{)}
    \PY{n+nv}{Cx} \PY{o}{=} \PY{n+nv}{arg2real}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{)}
    \PY{n+nv}{Cy} \PY{o}{=} \PY{n+nv}{arg2real}\PY{p}{(}\PY{l+m+mi}{4}\PY{p}{)}
    \PY{n+nv}{nt} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{5}\PY{p}{)}
    \PY{n+nv}{it} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{6}\PY{p}{)}
    \PY{n+nv}{dec} \PY{o}{=} \PY{n+nv}{arg2int}\PY{p}{(}\PY{l+m+mi}{9}\PY{p}{)}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{l+m+mi}{7}\PY{p}{,} \PY{n+nv}{fname}\PY{p}{)}

\PY{c}{!listing25}
    \PY{k}{call }\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{ctor}\PY{p}{(}\PY{n+nv}{it}\PY{p}{,} \PY{n+nv}{bcx}\PY{p}{,} \PY{n+nv}{bcy}\PY{p}{,} \PY{n+nv}{nx}\PY{p}{,} \PY{n+nv}{ny}\PY{p}{)}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} 
    \PY{k}{call }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{ptr}\PY{p}{)}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{courant}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{Cx}

    \PY{n+nv}{ptr} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{courant}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)} 
    \PY{n+nv}{ptr} \PY{o}{=} \PY{n+nv}{Cy}

    \PY{k}{call }\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{solve}\PY{p}{(}\PY{n+nv}{nt}\PY{p}{)}
\PY{c}{!listing26}

    \PY{k}{block}
\PY{k}{      }\PY{k+kt}{logical} \PY{k+kd}{::} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.FALSE.}
      \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{n+nv}{tmp}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
      \PY{k+kt}{real} \PY{k+kd}{::} \PY{n+nv}{diff}
      \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
      \PY{k}{allocate}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{(}\PY{n+nv}{nx}\PY{p}{,}\PY{n+nv}{ny}\PY{p}{)}\PY{p}{)}
      \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{l+m+mi}{8}\PY{p}{,} \PY{n+nv}{fname}\PY{p}{)}
      \PY{k}{call }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
      \PY{n+nv}{diff} \PY{o}{=} \PY{n+nb}{maxval}\PY{p}{(}\PY{n+nb}{abs}\PY{p}{(}\PY{n+nv}{slv}\PY{p}{\PYZpc{}}\PY{n+nv}{state}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{n+nv}{tmp}\PY{p}{)}\PY{p}{)}
      \PY{k}{print}\PY{o}{*}\PY{p}{,} \PY{l+s+s2}{"diff="}\PY{p}{,} \PY{n+nv}{diff}
      \PY{k}{if} \PY{p}{(}\PY{n+nv}{diff} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mf}{.5} \PY{o}{*} \PY{l+m+mi}{1}\PY{l+m+mf}{0.}\PY{o}{**}\PY{p}{(}\PY{o}{-}\PY{n+nv}{dec}\PY{p}{)}\PY{p}{)} \PY{n+nv}{error} \PY{o}{=} \PY{n+nb}{.TRUE.}
      \PY{k}{deallocate}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{)}
      \PY{k}{if} \PY{p}{(}\PY{n+nv}{error}\PY{p}{)} \PY{k}{then}
        \PY{c}{!print*, slv\PYZpc{}state()}
        \PY{k}{stop }\PY{l+m+mi}{1}
      \PY{k}{end }\PY{k}{if}
\PY{k}{    }\PY{k}{end }\PY{k}{block}
\PY{k}{  }\PY{k}{end }\PY{k}{block}

\PY{k}{  }\PY{k}{contains}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{arg2int}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{implicit }\PY{k}{none}
\PY{k}{    }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{argno}\PY{p}{,} \PY{k}{return}
\PY{k}{    }\PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{tmp}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
    \PY{k}{read}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{,}\PY{o}{*}\PY{p}{)}\PY{k}{return}
\PY{k}{  }\PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{function }\PY{n+nv}{arg2real}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{)} \PY{k}{result}\PY{p}{(}\PY{k}{return}\PY{p}{)}
    \PY{k}{implicit }\PY{k}{none}
\PY{k}{    }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{argno}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)} \PY{k+kd}{::} \PY{k}{return}
\PY{k}{    }\PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{l+m+mi}{666}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{tmp}
    \PY{k}{call }\PY{n+nb}{get\PYZus{}command\PYZus{}argument}\PY{p}{(}\PY{n+nv}{argno}\PY{p}{,} \PY{n+nv}{tmp}\PY{p}{)}
    \PY{k}{read}\PY{p}{(}\PY{n+nv}{tmp}\PY{p}{,}\PY{o}{*}\PY{p}{)}\PY{k}{return}
\PY{k}{  }\PY{k}{end }\PY{k}{function}

\PY{k}{  }\PY{k}{subroutine }\PY{n+nv}{read\PYZus{}file}\PY{p}{(}\PY{n+nv}{fname}\PY{p}{,} \PY{k}{array}\PY{p}{)}
    \PY{k+kt}{character} \PY{p}{(}\PY{n+nb}{len}\PY{o}{=}\PY{o}{*}\PY{p}{)}\PY{p}{,} \PY{k}{intent} \PY{p}{(}\PY{n+nv}{in}\PY{p}{)} \PY{k+kd}{::} \PY{n+nv}{fname}
    \PY{k+kt}{real}\PY{p}{(}\PY{n+nv}{real\PYZus{}t}\PY{p}{)}\PY{p}{,} \PY{k}{pointer} \PY{k+kd}{::} \PY{k}{array}\PY{p}{(}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{)}
    \PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{un}
    \PY{k}{open}\PY{p}{(}\PY{n+nv}{newunit}\PY{o}{=}\PY{n+nv}{un}\PY{p}{,} \PY{n+nv}{file}\PY{o}{=}\PY{n+nv}{fname}\PY{p}{,} \PY{n+nv}{status}\PY{o}{=}\PY{l+s+s1}{'old'}\PY{p}{,} \PY{n+nv}{action}\PY{o}{=}\PY{l+s+s1}{'read'} \PY{p}{)}
    \PY{k}{block    }
\PY{k}{      }\PY{k+kt}{integer} \PY{k+kd}{::} \PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}
      \PY{k}{do }\PY{n+nv}{i}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{k}{array}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
        \PY{k}{read}\PY{p}{(}\PY{n+nv}{un}\PY{p}{,} \PY{o}{*}\PY{p}{)} \PY{p}{(}\PY{k}{array} \PY{p}{(}\PY{n+nv}{i}\PY{p}{,} \PY{n+nv}{j}\PY{p}{)}\PY{p}{,} \PY{n+nv}{j}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n+nv}{size}\PY{p}{(}\PY{k}{array}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{)}
      \PY{k}{end }\PY{k}{do}
\PY{k}{    }\PY{k}{end }\PY{k}{block}
\PY{k}{    }\PY{k}{close} \PY{p}{(}\PY{n+nv}{un}\PY{p}{)}
   \PY{k}{end }\PY{k}{subroutine }\PY{n+nv}{read\PYZus{}file}
\PY{k}{end }\PY{k}{program}
\end{Verbatim}